---
title: "Using the fabPrediction Package for Nonparametric Prediction"
author: "Elizabeth Bersson"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(devtools)
```


The `fabPrediction` package provides functions to implement FAB conformal prediction algorithms as in [Bersson and Hoff, 2022](https://arxiv.org/pdf/2204.08122.pdf).


## Installation

```{r devtools, eval=FALSE}
devtools::install_github("betsybersson/fabPrediction")
library(fabPrediction)
```


## Demo


We will demonstrate FAB prediction with log radon values from counties in Minnesota. See [Price et al., 1996](https://pubmed.ncbi.nlm.nih.gov/8919076/). 

```{r data}
data(radon)
data(W)
```


### Single Population Analysis

If we have prior information regarding the population mean of the response, we can obtain FAB prediction regions for one county. 

```{r onecounty}
y_county9 = radon$radon[radon$group==9]
```


A FAB prediction region, although nonparametric, constructs prediciton intervals based on a standard normal-normal model,
\begin{aligned}
y \sim{}& N(\theta,\sigma^2)\\
\theta \sim{}& N(\mu,\tau^2\sigma^2),
\end{aligned}
and the algorithm requires input values of $\{\mu,\tau^2\}$. 

Say we are fairly confident log radon values in County 9 should be centered at 0.5. Then, we can obtain a prediction region with the `fab_prediction` function with a nonparametric guarantee of $1-\epsilon$ conservative coverage.


```{r fabcontinuous}
fab.region = predictionInterval(y_county9,method = "FAB",
                          epsilon = .15,
                          mu = 0.5,tau2 = 1)
fab.region$bounds
```

So, with at least $85\%$ probability, the next observed radon value in County 9 will be in the range contained in the object `fab.region`. We can easily plot the output:

```{r plotfab, fig.height=4, fig.width = 6,warning=F}
plot(fab.region,
     main="FAB",xlab="log(radon)")
```

We can easily compare this the distance-to-average conformal prediction region.

```{r plotdta, fig.height=4, fig.width = 6,warning=F}
plot(predictionInterval(y_county9,method = "DTA",epsilon = .15),
     main="DTA",xlab="log(radon)")
```

The functions include capabilities to obtain and plot standard frequentist and Bayesian prediction intervals based on a normal model. Unlike the FAB and DTA methods, the coverage of these methods rely on, among other things, accuracy of distributional assumptions.

```{r parametric,warning=F}
predictionInterval(y_county9,method = "direct",epsilon = .15)$bounds
predictionInterval(y_county9,method = "Bayes",epsilon = .15,
                   mu=0.5, tau2=1)$bounds
```


### Multiple Population Analysis

The package also features capability to handle multiple populations. Spatial relationships among the populations and covariates can be used to estimate conformal parameters. 

```{r, fig.height=4, fig.width = 6,warning=F}
params = EB_values(1,radon$radon,radon$group,W,X=rep(1,length(W)))
plot(predictionInterval(y,method = "FAB",epsilon = .15,
                        mu = params$mu, tau2 = params$tau2),
     main="FAB PI For County 1",xlab="log(radon)")
```





