---
title: "Using the fabPrediction Package to obtain Nonparametric Prediction Regions"
author: "Elizabeth Bersson"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(devtools)
library(fabPrediction)
```


The `fabPrediction` package provides functions to implement FAB conformal prediction algorithms for continuous data as in [Bersson and Hoff, 2022](https://arxiv.org/pdf/2204.08122.pdf)
and categorical counts data as in [Bersson and Hoff, 2023](https://arxiv.org/pdf/2311.15860.pdf).
The package also includes capabilities to obtain standard direct and Bayesian prediction regions for both types of data.


## Installation

```{r devtools, eval=FALSE}
devtools::install_github("betsybersson/fabPrediction")
library(fabPrediction)
```


## Demo - Countinuous Data


We will demonstrate FAB prediction with log radon values from counties in Minnesota. See [Price et al., 1996](https://pubmed.ncbi.nlm.nih.gov/8919076/). 

```{r data}
data(radon)
data(W)
```


### Single Group Analysis

If we have prior information regarding the population mean of the response, we can obtain FAB prediction regions for one county. 

```{r onecounty}
y_county9 = radon$radon[radon$group==9]
```


A FAB prediction region, although nonparametric, constructs prediciton intervals based on a standard normal-normal model,
\begin{aligned}
y \sim{}& N(\theta,\sigma^2)\\
\theta \sim{}& N(\mu,\tau^2\sigma^2),
\end{aligned}
and the algorithm requires input values of $\{\mu,\tau^2\}$. 

Say we are fairly confident log radon values in County 9 should be centered at 0.5. Then, we can obtain a prediction region with the `fab_prediction` function with a nonparametric guarantee of $1-\alpha$ conservative coverage.


```{r fabcontinuous}
fab.region = predictionInterval(y_county9,method = "FAB",
                          alpha = .15,
                          mu = 0.5,tau2 = 1)
fab.region$bounds
```

So, with at least $85\%$ probability, the next observed radon value in County 9 will be in the range contained in the object `fab.region`. We can easily plot the output:

```{r plotfab, fig.height=4, fig.width = 6,warning=F}
plot(fab.region,
     main="FAB",xlab="log(radon)")
```

We can easily compare this the distance-to-average conformal prediction region.

```{r plotdta, fig.height=4, fig.width = 6,warning=F}
plot(predictionInterval(y_county9,method = "DTA",alpha = .15),
     main="DTA",xlab="log(radon)")
```

The functions include capabilities to obtain and plot standard frequentist and Bayesian prediction intervals based on a normal model. Unlike the FAB and DTA methods, the coverage of these methods rely on, among other things, accuracy of distributional assumptions.

```{r parametric,warning=F}
predictionInterval(y_county9,method = "direct",alpha = .15)$bounds
predictionInterval(y_county9,method = "Bayes",alpha = .15,
                   mu=0.5, tau2=1)$bounds
```


### Multiple Group Analysis

The package also features capability to handle multiple groups. Spatial relationships among the groups and covariates can be used to estimate conformal parameters. 

```{r, fig.height=4, fig.width = 6,warning=F}
params = EB_values(9,radon$radon,radon$group,W,X=rep(1,nrow(W)))
plot(predictionInterval(y_county9,method = "FAB",alpha = .15,
                        mu = params$mu, tau2 = params$tau2),
     main="FAB PI For County 9", xlab="log(radon)")
```



## Demo - Categorical Data

We will demonstrate constructing prediction sets for categorical data on a simulated dataset consisting of 50 categories and 5 groups with varying within-group sample sizes.

```{r catdata}
N.groups = c(10,50,75,100,150)

set.seed(1)
prob = rdirichlet(50:1)
y = t(sapply(N.groups,function(j)rmultinom(1,j,prob)))
```


### Single Group Analysis

If we have prior information regarding the prior concentration parameter for a given group, we can construct a FAB prediction set for that group.

```{r}
y_group3 = y[3,]
fab.set = predictionSet(y_group3,method = "FAB",
                        gamma = c(50:1))
```

And, we can plot the categories included in the prediction set:

```{r}
plot(fab.set, main = "FAB Prediction Set for Group 3",
     cex.axis=.5)
```

Similarly, we can construct and plot a direct prediction set for the same data:

```{r}
plot(predictionSet(y_group3,method = "direct"), 
     main = "Direct Prediction Set for Group 3",
     cex.axis=.5)
```

We can also construct a Bayesian prediction set, using the same prior as before in the FAB approach. Note, though, that this prediction set does not necessarily have the guaranteed coverage rate.

```{r}
plot(predictionSet(y_group3,method = "Bayes",gamma = c(50:1)), 
     main = "Bayes Prediction Set for Group 3",
     cex.axis=.5)
```



### Multiple Group Analysis

The package also features capability to handle multiple groups. To construct a prediction set for group 3 in this way, the prior concentration $\gamma$ can be learned from all groups except 3. 

```{r}
gamma0 = polyaMLE(y[-3,])
```

This prior can be used as the prior information in constructing a FAB prediction set.

```{r}
plot(predictionSet(y_group3,method = "FAB",gamma = gamma0), 
     main = "FAB Prediction Set for Group 3 using Indirect Information",
     cex.axis=.5)
```


